---
title: "programming_in_R"
format: pdf
editor: visual
---

## Task One: Conceptual Questions

### Question 1: What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

The `lapply()` function is used for list objects. It applies the specified function to every element of the list and then returns a list. The equivalent `purrr` function is `map()`.

### Question 2: Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns
are numeric). We want use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")`
on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify
`method = "kendall"` when calling `lapply()`)

```{r}
lapply(X = my_list, 
       FUN = cor,
       method = "kendall")
```

### Question 3: What are two advantages of using `purrr` functions instead of the `BaseR` apply family?

One advantages of the `purrr` functions is that they are generally more consistent (additional arguments can be specified under the same naming conventions, the order of arguments is consistent, etc.). Another advantage is that the `purrr` functions contain useful helper functions (like `partial()`) and variants (like `map2()`). 

### Question 4: What is a side-effect function?

A side-effect function uses the data to perform a task, but when the task is complete it returns the original data unmodified. 

### Question 5: Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

Variables that are created in functions are only stored temporarily in the function environment. They are not kept after the function call is finished, and therefore never stored in the global environment. 

## Task Two: Writing R Functions

### Question 1: Writing RMSE function

```{r}
getRMSE <- function(resp, pred, ...) {
  sq_diff <- (resp - pred)^2
  mean <- mean(sq_diff, ...)
  rmse <- sqrt(mean)
  return(rmse)
  }
```

### Question 2: Testing RMSE function

```{r}
#Generating test data
set.seed(10) #setting random seed so that data can be replicated
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
#Testing my RMSE function
getRMSE(resp, pred)
```

```{r}
#Replacing several observations with missing values
resp[3] <- NA
resp[55] <- NA
```

```{r}
#Testing my RMSE function without specifying missing values behavior
getRMSE(resp, pred)
```

```{r}
#Testing my RMSE function with specifying missing values behavior
getRMSE(resp, pred, na.rm = TRUE)
```

## Question 3: Writing Mean Absolute Deviation Function

```{r}
getMAE <- function(resp, pred, ...){
  diff <- resp - pred
  abs_diff <- abs(diff)
  mae <- mean(abs_diff, ...)
  return(mae)
}
```

## Question 4: Testing MAE Function

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
#Testing my MAE function
getMAE(resp, pred)
```

```{r}
#Replacing several observations with missing values
resp[3] <- NA
resp[55] <- NA
```

```{r}
#Testing my MAE function without specifying missing values behavior
getMAE(resp, pred)
```

```{r}
#Testing my MAE function with specifying missing values behavior
getMAE(resp, pred, na.rm = TRUE)
```

### Question 5: Writing a Wrapper Function to Calculate RMSE and MAE

```{r}
wrapper <- function(resp, pred, metric = "both", ...) {
  if (!is.vector(resp) | !is.numeric(resp) | !is.vector(pred) | !is.numeric(pred)) {
    print("ERROR: the responses and/or the predictions are not a numeric vector")
    stop()
  } else if (metric == "both") {
    rmse <- getRMSE(resp, pred, ...)
    mae <- getMAE(resp, pred, ...)
    return(list("RMSE" = rmse, "MAE" = mae))
  } else if (metric == "rmse" | metric == "RMSE") {
    rmse <- getRMSE(resp, pred, ...)
    return("RMSE" = rmse)
  } else if (metric == "mae" | metric == "MAE") {
    mae <- getMAE(resp, pred, ...)
    return("MAE" = mae)
  }
  return(list("rmse" = rmse, "mae" = mae))
}
```

### Question 6: Testing the Wrapper Function

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
wrapper(resp, pred)
```

```{r}
wrapper(resp, pred, "rmse")
```

```{r}
wrapper(resp, pred, "mae")
```

```{r}
#Replacing several observations with missing values
resp[3] <- NA
resp[55] <- NA
```

```{r}
wrapper(resp, pred, na.rm = TRUE)
```

```{r}
wrapper(resp, pred, "rmse", na.rm = TRUE)
```

```{r}
wrapper(resp, pred, "mae", na.rm = TRUE)
```

```{r}
wrapper(resp, cars)
```

